---
title: "Fitness Data Analysis"
author: "Shahmir Zubair"
date: "2023-02-02"
output:
  word_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

## Scenario

You are a junior data analyst working on the marketing analyst team at
Bellabeat, a high-tech manufacturer of health-focused products for
women. Bellabeat is a successful small company, but they have the
potential to become a larger player in the

global smart device market. Urška Sršen, cofounder and Chief Creative
Oﬃcer of Bellabeat, believes that analyzing smart device ﬁtness data
could help unlock new growth opportunities for the company. You have
been asked to focus on one of Bellabeat's products and analyze smart
device data to gain insight into how consumers are using their smart
devices. The insights you discover will then help guide marketing
strategy for the company. You will present your analysis to the
Bellabeat executive team along with your high-level recommendations for
Bellabeat's marketing strategy.

## Characters and products

### ● Characters

○ Urška Sršen: Bellabeat's cofounder and Chief Creative Oﬃcer

○ Sando Mur: Mathematician and Bellabeat's cofounder; key member of the
Bellabeat executive team

○ Bellabeat marketing analytics team: A team of data analysts
responsible for collecting, analyzing, and reporting data that helps
guide Bellabeat's marketing strategy. You joined this team six months
ago and have been busy learning about Bellabeat''s mission and business
goals --- as well as how you, as a junior data analyst, can help
Bellabeat achieve them.

### ● Products

○ Bellabeat app: The Bellabeat app provides users with health data
related to their activity, sleep, stress, menstrual cycle, and
mindfulness habits. This data can help users better understand their
current habits and make healthy decisions. The Bellabeat app connects to
their line of smart wellness products.

○ Leaf: Bellabeat's classic wellness tracker can be worn as a bracelet,
necklace, or clip. The Leaf tracker connects to the Bellabeat app to
track activity, sleep, and stress.

○ Time: This wellness watch combines the timeless look of a classic
timepiece with smart technology to track user activity, sleep, and
stress. The Time watch connects to the Bellabeat app to provide you with
insights into your daily wellness.

○ Spring: This is a water bottle that tracks daily water intake using
smart technology to ensure that you are appropriately hydrated
throughout the day. The Spring bottle connects to the Bellabeat app to
track your hydration levels.

○ Bellabeat membership: Bellabeat also oﬀers a subscription-based
membership program for users. Membership gives users 24/7 access to
fully personalized guidance on nutrition, activity, sleep, health and
beauty, and mindfulness based on their lifestyle and goals.

## About the company

Urška Sršen and Sando Mur founded Bellabeat, a high-tech company that
manufactures health-focused smart products. Sršen used her background as
an artist to develop beautifully designed technology that informs and
inspires women around the world. Collecting data on activity, sleep,
stress, and reproductive health has allowed Bellabeat to empower women
with knowledge about their own health and habits. Since it was founded
in 2013, Bellabeat has grown rapidly and quickly positioned itself as a
tech-driven wellness company for women.

By 2016, Bellabeat had opened oﬃces around the world and launched
multiple products. Bellabeat products became available through a growing
number of online retailers in addition to their own e-commerce channel
on their website. The company has invested in traditional advertising
media, such as radio, out-of-home billboards, print, and television, but
focuses on digital marketing extensively. Bellabeat invests year-round
in Google Search, maintaining active Facebook and Instagram pages, and
consistently engages consumers on Twitter. Additionally, Bellabeat runs
video ads on Youtube and display ads on the Google Display Network to
support campaigns around key marketing dates.

Sršen knows that an analysis of Bellabeat's available consumer data
would reveal more opportunities for growth. She has asked the marketing
analytics team to focus on a Bellabeat product and analyze smart device
usage data in order to gain insight into how people are already using
their smart devices. Then, using this information, she would like
high-level recommendations for how these trends can inform Bellabeat
marketing strategy.

## Ask

Sršen asks you to analyze smart device usage data in order to gain
insight into how consumers use non-Bellabeat smart devices. She then
wants you to select one Bellabeat product to apply these insights to in
your presentation. These questions will guide your analysis:

1\. What are some trends in smart device usage?

2\. How could these trends apply to Bellabeat customers?

3\. How could these trends help inﬂuence Bellabeat marketing strategy?

## Analysis

First, we will install and load some important packages.

```{r warning=FALSE, include=FALSE}
# Loading packages
library(tidyverse)
library(ggplot2)
```

Reading all the necessary csv files into program

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
daily_activity <- read_csv('Data/dailyActivity_merged.csv')

weight_log <- read_csv('Data/weightLogInfo_merged.csv')

sleep_day <- read_csv('Data/sleepDay_merged.csv')
```

### Correlation between exercise and sedentary

Data shows a clear trend between Total steps and sedentary minutes. It
suggests that people who exercise more spend less time in sedentary as
opposed to people who exercise less.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

ggplot(data=daily_activity, aes(x=TotalSteps, y=SedentaryMinutes)) +
geom_point(color = "#004F98") +
geom_smooth(method=lm)

```

### Correlation between exercise and calories burned

Data also proves that more exercise leads to more calories burned.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
  
ggplot(daily_activity, aes(TotalSteps, Calories)) +
  geom_point(color = "#004F98") +
  geom_smooth(method=lm)
  
```

### Correlation between excercise and weight

Grouping and summarizing the data.

```{r message=TRUE, warning=FALSE}
steps_vs_weight <- merge(daily_activity, weight_log) %>% 
  group_by(Id) %>% 
  summarize(steps = mean(TotalSteps), weight = mean(WeightKg))

steps_vs_weight
```

There is a clear downwards trend between weight and steps. This suggests
that people who exercise more tend to have less weight.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(steps_vs_weight, aes(steps,weight)) + 
  geom_point(color="#004F98", size = 3) + 
  geom_smooth(method = lm, level = 0.4, color = "#004F98")

```

### Correlation between very active people and sleep (No Relation found)

Grouping and summarizing the data

```{r}
activity_vs_sleep <- merge(daily_activity, sleep_day) %>%
  select(Id, VeryActiveMinutes, TotalMinutesAsleep) %>% 
  group_by(Id) %>% 
  summarize(very_active_minutes = mean(VeryActiveMinutes), total_minutes_asleep = mean(TotalMinutesAsleep))

activity_vs_sleep
```

The plot suggests that sleep is not effected by being active in the day.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(activity_vs_sleep, aes(very_active_minutes, total_minutes_asleep)) +
  geom_point(color = "#004F98", size = 3) +
  geom_smooth(method=lm)
```

### Correlation between VeryActiveMinutes vs day of the week

Making a dayname column in daily_activity dataframe. Since ActivityDate
is char datatype, first i will convert it into date.

```{r}
daily_activity$ActivityDate <- as.Date(daily_activity$ActivityDate,"%m/%d/%y")
daily_activity$dayname <- weekdays(daily_activity$ActivityDate)
```

There is a clear downwards trend between being very active and day of
the week. This may be due to high motivation levels at the start of the
week which gradually go down as the week progresses.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(daily_activity, aes( x = fct_relevel(dayname, "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), VeryActiveMinutes)) + 
  geom_col(fill = "#004F98") + 
  labs(x="DayName") 
```

### How many people are overweight? (using BMI)

Creating a new dataframe for 8 people who submitted their weight
information

```{r}
BMI <- weight_log %>% 
  group_by(Id) %>% 
  summarise(mean_BMI = mean(BMI, na.rm = TRUE))
```

Creating a column for their health condition based on BMI (Body Mass
Index).

***Note: BMI less than 25 is healthy. BMI between 25-25.9 is overweight
and greater than 30 is obese***

```{r}
BMI <- BMI %>%  mutate(case_when(
  mean_BMI < 25 ~ "Healthy",
  mean_BMI >=25 & mean_BMI < 30 ~ "Overweight",
  mean_BMI > 30 ~ "Obese"
)) 
colnames(BMI)[3] <- "health_condition"


```

Out of the 8 people only 3 are Healthy, 4 are overweight and only 1 is
obese.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(BMI, aes(fct_relevel(health_condition, "Obese", "Overweight", "Healthy"))) +
  geom_bar(fill = "#004F98") +
  labs(title = "Condition of 8 people based on BMI",x = "Condition", y = "Number of people")
```

## Answering the buisness question

1\. What are some trends in smart device usage?

2\. How could these trends apply to Bellabeat customers?(I believe that
displaying the trends would already indicate how Bellabeat customers
would follow suit.)

Based on insights generated by my analysis, i came to following
hypothesis:

1.  People spend more time engaged in physical activity on Sundays,
    which proceeds to decrease throughout the week and rise again on
    weekends. This may be due to motivation levels being higher on the
    weekend and then as the week progresses, motivation levels decrease
    as work load increases.

Now the Final Question.

3.  How could these trends help inﬂuence Bellabeat marketing strategy?

1\. Bellabeat could host events limited to Bellabeat's members which
reward users who engage in a healthy lifestyle(I.E 8 hours of sleep,8k
steps a day, less than 7 hours sedentary etc.) with credits. With enough
credits, users could then use credits to purchase Bellabeat's products
that help supplement a healthy lifestyle.

2\. Bellabeat could partner with brands (IE wellness, sports, health) to
reward users who consistently engage in a healthy lifestyle with
coupons/store discounts.

3\. With the 2 previous points combined, Bellabeat could select
previously unhealthy individuals (who are now healthy), interview them
and publish motivational videos as to how Bellabeat encouraged them to
have a change in lifestyle.

## **Thank You!**
